Learning Discrete-valued Bayesian Networks from Mixed Data

### 2015-11-20 BN_discretizer_paper.tex
misc

\newcommand{\bigo}{\mathcal{O}} % NOTE(tim): if you prefer normal O change it here.

@@

MODL discretization \citep{Boulle_2006} is a Bayesian method for discretizing a continuous feature according to a class variable.
Bayesian methods maximize the probability of the model given the data, $P(\textit{Model} \mid \textit{Data})$.
An application of Bayes' rule and constant $P(\textit{Data})$ shows that this is equivalent to maximizing

\begin{align}
P(\textit{Model}) \cdot P(\textit{Data} \mid \textit{Model})\text{,}
\end{align}

\noindent
where $P(\textit{Model})$ is a prior over models and ${P(\textit{Data} \mid \textit{Model})}$ is the data likelihood.
The data likelihood usually increases with the number of discretization intervals, as more intervals allow for richer representations.
The prior is chosen to decrease as the number of discretization intervals increases to favor simpler models and prevent overfitting.
Maximizing ${P(\textit{Model}) \cdot P(\textit{Data} \mid \textit{Model})}$ requires a trade-off when determining the number of discretization intervals.

@@

Let $\cX$ be a continuous variable and let $\cx$ be a specific instance of $\cX$.
A discretization policy $\Lambda_{\cX} = \curly{e_1 < e_2 < \ldots < e_{k-1}}$ for $\cX$ is a mapping from $\mathbb{R}$ to $\curly{1,2,3,\ldots,k}$ such that

@@

The discretization edges $e_{1:k-1}$ mark the boundaries between discretization intervals.
In this work they are restricted to the midpoints between unique ascending instances of $\cX$.
Thus, each edge $e_i = \paren{\cu_{j} + \cu_{{j+1}}}/2$ for some $j$.
Two useful integer representation of $\Lambda_{\cX}$ can be written

\begin{equation}
\label{eq:disc_def}
  \Lambda_{\cX} = \curly{\lambda_1 < \lambda_2  < \ldots < \lambda_k} \equiv \curly{\gamma_1 , \gamma_2  , \ldots , \gamma_k}\text{,}
\end{equation}

@@

This section describes experiments conducted to evaluate the proposed methods.
All experiments were run on datasets from the publically available University of California, Irvine machine learning repository \citep{Lichman_2013}.
Variables are labelled alphabetically in the order given on the dataset information webpage
In the figures that follow, shaded nodes correspond to initially discrete variables and the subscripts indicate the number of discrete instantiations.

Two experiments were conducted on each dataset.
The first experiment compares the performance of the optimal Bayesian discretization method, the heuristic Bayesian discretization method, and MDL on a known Bayesian network structure.
The structure was obtained by prediscretizing each continuous variable into $k$ uniform-width intervals, where $k$ is the median number of instantiations of the discrete variables, and using the structure with the highest Bayesian score from one thousand runs of the K2 algorithm with random variable orderings.
The second experiment compares the same methods applied when simultaneously discretizing and learning network structure.

The discretizations are compared using the mean cross validated log-likelihood of the data $D$ given the graph structure $G$ and discretization policies $\discset$.
The log-likelihood has two components,

@@

The optimal discretization policy for MDL is found with dynamic programming in a similar manner as with the Bayesian method.
For a Bayesian network with a single continuous variable $\cX$, the MDL objective function is

@@

The Bayesian approaches and the MDL discretization method were tested on the Auto MPG data using the network network shown in Figure~\ref{fig:auto_graph_1}.
This structure was obtained by prediscretizing each continuous variable into five uniform-width intervals, where five is the median cardinality of the discrete variables, and then taking the structure with highest likelihood from one thousand runs of K2.

@@

Table~\ref{table:auto_disc_table_1} lists the discretization edges and mean log-likelihoods under 10-fold cross validation of the discrete Bayesian network resulting from each discretization method.
The results indicate that the heuristic Bayesian method produces exactly the same discretization policy as the optimal Bayesian method.
The MDL method does not produce any discretization edges, assigning one continuous interval to each continuous variable, and produces the structure with the lowest score.
The cause behind MDL's discretization policy is discussed in Section~\ref{subsec:discuss_exp}.

\begin{table}[h]
  \centering
  \caption{
    Results from discretization of the Auto MPG dataset with fixed structure from Figure~\ref{fig:auto_graph_1}.
    The first five rows list the discretization edges for each continuous variable.
    The last row lists the mean cross-validated log-likelihood; positive values are better.
    The optimal and heuristic Bayesian discretization policies are identical and the MDL method does not produce any discretization edges.
  }
  \input{table_autoMPG_1.tex}
  \label{table:auto_disc_table_1}
\end{table}

Figure~\ref{fig:auto_exp1_distr_1_3} compares the Bayesian discretizaton policy for variables $A$ and $C$ with the original Auto MPG data.
The black marks are the original Auto MPG samples.
The color of a discretized region indicates the marginal probability of a sample from $P(A,C)$ being drawn from that region.
As shown in the figure, the discretization edges obtained by the optimal Bayesian approach and the resulting marginalized probability match the raw data well.

\begin{figure}[ht]
\centering
  \input{graph_auto_exp1_distr_1_3}
  \caption{
    Comparison of the Bayesian discretization policy for variables $A$ and $C$ to the original Auto MPG data learned with a fixed network.
    The marginal probabilities of discretized regions closely match the raw data.
  }
  \label{fig:auto_exp1_distr_1_3}
\end{figure}

@@

In this experiment the network structure was not fixed in advance and was learned simultaneously with the discretization policies.
Figure~\ref{fig:auto_graph_2} shows a learned Bayesian network structure and the corresponding numbers of intervals after discretization for each continuous variable.
This result was obtained by running Algorithm~\ref{alg:structure_learn} fifty times using the optimal Bayesian method and choosing the structure with the highest likelihood.

\begin{figure}[ht]
  \centering
  \input{graph_autocar_2}
  \caption{The discrete-valued Bayesian network learned from the Auto MPG dataset}
  \label{fig:auto_graph_2}
\end{figure}

Figure~\ref{fig:auto_exp2_distr_1_3} compares the Bayesian discretizaton policy for variables $A$ and $C$ in the learned network with the original Auto MPG data.
The color of a discretized region indicates the marginal probability of a sample from $P(A,C)$ being drawn from that region.
Although there are fewer discretization edges for $A$ and $C$ in the learned network, the marginal distribution is still captured.
Although the discretization result of the optimal Bayesian approach may vary according to the network structure, it still produces high-quality discretizations.

\begin{figure}[ht]
  \centering
  \input{graph_auto_exp2_distr_1_3}
  \caption{
    Comparison of the optimal Bayesian discretization policy for variables $A$ and $C$ to the original Auto MPG data learned along with the network structure.
    Although the number of discretization edges is less than those in Figure~\ref{fig:auto_exp1_distr_1_3}, the probability distribution still closely matches the original data.
  }
  \label{fig:auto_exp2_distr_1_3}
\end{figure}

@@

The Bayesian approaches and the MDL discretization method were tested on the Wine data using the network network shown in Figure~\ref{fig:wine_graph_1}.
This structure was obtained by prediscretizing each continuous variable into three uniform-width intervals, where three is the median cardinality of the discrete variables, and then taking the structure with highest likelihood from one thousand runs of K2.

Table~\ref{table:wine_disc_table_1} lists the discretization edges and mean log-likelihoods under 10-fold cross validation of the discrete Bayesian network resulting from each discretization method.
The Bayesian methods outperform the MDL method in likelihood by a significant margin.
The heuristic method discretizes the variables $G$, $H$, and $J$ into one fewer interval and the variable $M$ into one more interval than the optimal Bayesian approach, resulting in a slightly lower mean cross validated log likelihood.
The overall performance of the heuristic method is still very high, and if runtime is the primary concern, the heuristic method can be considered the better approach.
\tim{Do we need to list runtime too?}

Notice that there are some discretization edges which appear in all three discretization methods, such as \num{1.42} and \num{2.35} for variable $C$ and \num{0.785} for variable $L$.
This indicates that MDL indeed can find some important discretization edges, but it is not sensitive enough to find more edges.

\begin{figure}[ht]
  \centering
  \scalebox{0.8}{\input{graph_wine_1}}
  \caption{Structure from prediscretized K2 structure learning to compare the methods for Wine dataset}
  \label{fig:wine_graph_1}
\end{figure}

\begin{table}
  \centering
  \caption{
    Results from discretization of the Wine MPG dataset with fixed structure from Figure~\ref{fig:wine_graph_1}.
    The first thirteen rows list the discretization edges for each continuous variable.
    The last row lists the mean cross-validated log-likelihood; positive values are better.
    The optimal Bayesian approach has the best discretization result in terms of likelihood, but the heuristic method obtains a similar result with nearly the same likelihood.
    The MDL method produces discretization edges for more than half of the continuous variables but does not produce enough discretization intervals.
    Some discretization edges appear in the results of all three methods.
  }
  \scalebox{1.0}{
  \input{table_wine_1.tex}
  }
  \label{table:wine_disc_table_1}
\end{table}

Figure~\ref{fig:wine_exp1_distr} compares the optimal Bayesian and MODL discretizaton policies for variables $E$ and $K$ with the original Wine data.
The discretization edges \num{17.9} for $E$ and \num{34.6} for $K$ appear in both plots.
The MDL method does not use enough intervals for discretization.
Relative sensitivities of each method to the input data is discussed in Section~\ref{subsec:discuss_exp}.

\begin{figure}[ht]
  \centering
  \input{graph_wine_exp1_distr}
  \caption{Comparison of the discretization policies for variables $A$ and $C$ obtained using optimal Bayesian and MDL methods}
  \label{fig:wine_exp1_distr}
\end{figure}

@@

Figure~\ref{fig:wine_exp2_distr} shows the discretization policy for variables $E$ and $K$ obtained with the optimal Bayesian method.
The discretization edge $E = 17.9$ also appears in all discretization policies from the fixed network (Figure~\ref{fig:wine_exp1_distr}).
This indicates that some discretization edges are robust against the network structure.
Furthermore, the discretization edge at $E = 23.5$ present in the fixed-structure case is missing in the learned structure.
This is caused by $E$ having twice as many parents in the learned network structure.
The more parents a variable has, the less discretization intervals it will be able to support, as the set of statistics required to define the resulting distribution increases exponentially with the number of parents.

\begin{figure}[ht]
  \centering
  \scalebox{0.7}{\input{graph_wine_2}}
  \caption{The discrete-valued Bayesian network learned from the Wine dataset using the optimal Bayesian method.}
  \label{fig:wine_graph_2}
\end{figure}

\begin{figure}[ht]
  \centering
  \input{graph_wine_exp2_distr_5_11}
  \caption{
    The discretization policy for variables $E$ and $K$ on the learned network in Figure~\ref{fig:wine_graph_2}.
    The discretization edge $E = 17.9$ also appears in the policies for the fixed network.
    This indicates that some discretization edges are robust against changing network structure.
  }
  \label{fig:wine_exp2_distr}
\end{figure}

@@

The Housing dataset contains variables related to the values of houses in Boston suburbs.
The dataset has \num{506} samples over fourteen variables.
Only variables $D$ and $I$ are discrete-valued, with \num{2} and \num{9} instantiations, respectively.
Variable $A$ is the only discrete variable and has three instantiations.
Despite their being continuous, several variables in the Housing dataset possess many repeated values.
The following experiments were conducted with an upper limit of three on the number of parents per variable to prevent memory issues with MDL.

@@

The Bayesian network structure in Figure~\ref{fig:housing_graph_1} was obtained by prediscretizing each continuous variable into five uniform-width interval and then running the K2 algorithm one thousand times and choosing the network with the highest likelihood.
Table~\ref{table:housing_disc_table_1} shows the numbers of intervals after discretization for each continuous variables and the log-likelihood of the dataset based on each discretization method.
MDL method does not produce any discretization edges for most variables.
The heuristic Bayesian method produces far more discretization edges than the optimal Bayesian method due to repeated instances in the dataset.
These numbers of intervals returned by the three methods illustrate the over-sensitive nature of the heuristic method and the under-sensitive nature of the MDL method.
\tim{I am not sure we are using `sensitive' correctly. Sensitive to what? Perhaps `biased towards higher edge counts'?}
The relative sensitivity of each algorithm will be discussed in Section~\ref{subsec:discuss_exp}.

The heurstic method produces the discretization with the highest cross-validated log likelihood, but although the heuristic method leads to a higher number of intervals, its likelihood is not significantly better than the optimal Bayesian approach.
This shows that the optimal Bayesian approach captures the most important discretization edges.

\begin{figure}[ht]
  \centering
  \scalebox{0.8}{\input{graph_housing_1}}
  \caption{Structure from prediscretized K2 structure learning to compare the methods for Housing dataset}
  \label{fig:housing_graph_1}
\end{figure}

\begin{table}
  \centering
  \caption{
    Discretization policy summary of the Housing dataset based on the fixed network structure shown in Figure~\ref{fig:housing_graph_1}.
    The first twelve rows show the numbers of intervals after discretization by each method.
    The last row is the mean cross-validated log-likelihood of the dataset based on the discretization results.
    The heuristic method is over-sensitive \todo{to what?} and discretizes some variables into too many intervals.
    The MDL method is less sensitive and on many cases does not produce any discretization intervals.
    The optimal Bayesian approach produces a reasonable number of discretization edges.
  }
  \scalebox{1.0}{\input{table_housing_1.tex}}
  \label{table:housing_disc_table_1}
\end{table}

Figures~\ref{fig:housing_exp1_distr_3_5} and \ref{fig:housing_exp1_distr_8_5} show the discretization policy learned using the optimal Bayesian approach on the fixed network shown in Figure~\ref{fig:housing_graph_1}.
In Figure~\ref{fig:housing_exp1_distr_3_5}, the scatter points corresponding to the original dataset have been jittered, because there are many repeated values for variables $C$ and $E$.
Each repeated point forms a single discrete region, thereby encouraging discretization.
In contrast, the samples for $H$ are well spread out, resulting in fewer discretization regions and larger discretization intervals.

\begin{figure}[ht]
  \centering
  \input{graph_housing_exp1_distr_3_5}
  \caption{
    The discretization policies for variables $C$ and $E$ learned using the optimal Bayesian approach on the fixed network shown in Figure~\ref{fig:housing_graph_1}.
    The scatter points were jittered to reveal the repeated values in the dataset.
    The repeated values lead to a high number of discretization edges, because repeated values can easily form discretized regions by their own.
  }
  \label{fig:housing_exp1_distr_3_5}
\end{figure}

@@

Figure~\ref{fig:housing_graph_2} shows a learned Bayesian network structure and the corresponding numbers of intervals after discretization for each continuous variable.
Variable $D$ is neighborless is both the learned and fixed networks.
The continuous variables $C$, $E$, $J$ and $K$, which are all connected through $C$, have many discretization intervals.
Typically, when discretizing a variable, the expected number of discretization intervals is close to the highest cardinality among variables in its Markov blanket.
This naturally leads to clusters of variables with high numbers of discretization intervals.

@@

\begin{figure}[ht]
  \centering
  \input{graph_housing_exp2_distr_3_5}
  \caption{
    The discretization policy for variables $C$ and $E$ on the learned network in Figure~\ref{fig:housing_graph_2}.
    The high number of discretization edges are caused by the repeated values in variables $C$ and $E$.
  }
  \label{fig:housing_exp2_distr_3_5}
\end{figure}

@@

\begin{figure}[ht]
\centering
  \input{graph_exp_discuss}
  \caption{A simple Bayesian network that is used to demonstrate the sensitivity of the three methods}
  \label{fig:exp_discuss}
\end{figure}

To quantatively discuss the sensitivity of the three methods to the number of discretization intervals, consider the Bayesian network in Figure~\ref{fig:exp_discuss}, where $\cX$ is continuous and $P_1$ and $P_2$ are discrete.
If $\Pa_{\cX} = \curly{P_1, P_2}$ and $D_{\cX} = \curly{1.0,2.0,3.0, \ldots n}$, then the corresponding objective functions for the three discretization methods are:

\begin{small}
  \begin{equation}
  \begin{aligned}
  f_{\text{MDL}} & = \overbrace{Z_1 \cdot k + \ln(k) + \ln {{n +k - 1}\choose{k -1}} }^{\text{Penalty Term}}+   \overbrace{\rule{0pt}{1.35em} \ln(N) \cdot I(X^*, \Pa_X)}^{\text{Edge Position Term}} \\
  f_{\text{Bayesian}} &= \overbrace{\rule{0pt}{1.85em}  Z_2 \cdot k + \sum_{i=1}^k \ln {{\gamma_i + J_P - 1}\choose{J_P -1}}}^{\text{Penalty Term}} + \overbrace{\sum^k_{i=1} \ln \left(   {{\gamma_i !}\over{n^{(P)}_{i,1}! n^{(P)}_{i,2}! \cdots n^{(P)}_{i,J_P}!    }} \right) }^{\text{Edge Position Term}}\\
  f_{\text{Heuristic}} &= \overbrace{Z_2 \cdot k + \sum_{i=1}^k \ln {{\gamma_i + J_{P_1} - 1}\choose{J_{P_1} -1}} + \sum_{i=1}^k \ln {{\gamma_i + J_{P_2} - 1}\choose{J_{P_2} -1}} }^{\text{Penalty Term}}+\\
  & \overbrace{\sum^k_{i=1}  \left[ \ln \left(   {{\gamma_i !}\over{n^{(P_1)}_{i,1}! n^{(P_1)}_{i,2}! \cdots n^{(P_1)}_{i,J_{P_1}}    }} \right)  + \ln \left(   {{\gamma_i !}\over{n^{(P_2)}_{i,1}! n^{(P_2)}_{i,2} \cdots n^{(P_2)}_{i,J_{P_2}} }} \right) \right]}^{\text{Edge Position Term}}
  \end{aligned}\text{,}
  \end{equation}
\end{small}

\noindent
where $Z_1$ and $Z_2$ are constant over discretizations, $k$ is the number of discretization intervals, and $I\paren{A, B} = \sum_{a,b} \hat{P}(a,b) \ln\frac{\hat{P}(a,b)}{\hat{P}(a)\hat{P}(b)}$ is the mutual information based on estimated probability1ies.
Note that the third term of $f_\text{MDL}$ had been approximated using $H(p)$ (see Equation~\ref{eqn:MDL}) in the original work by \cite{Friedman_1996} but here is written without that approximation.

@@

\section*{Appendix}

>>>>

\newcommand{\bigo}{O} % NOTE(tim): if you prefer normal O change it here.

@@

MODL discretization \citep{Boulle_2006} is a Bayesian method for discretizing a continuous feature according to a class variable.
Bayesian methods maximize the probability of the model given the data, $P(\textit{Model} \mid \textit{Data})$.
An application of Bayes' rule and constant $P(\textit{Data})$ shows that this is equivalent to maximizing ${P(\textit{Model}) \cdot P(\textit{Data} \mid \textit{Model})}$, where $P(\textit{Model})$ is a prior over models and ${P(\textit{Data} \mid \textit{Model})}$ is the data likelihood.
The data likelihood usually increases with the number of discretization intervals, as more intervals allow for richer representations.
The prior is chosen to decrease as the number of discretization intervals increases to favor simpler models and prevent overfitting.
Maximizing ${P(\textit{Model}) \cdot P(\textit{Data} \mid \textit{Model})}$ requires a trade-off when determining the number of discretization intervals.

@@

Let $\cX$ be a continuous variable and let $\cx$ be a specific instance of $\cX$.
A discretization policy $\Lambda_{\cX} = \anglebrackets{e_1 < e_2 < \ldots < e_{k-1}}$ for $\cX$ is a mapping from $\mathbb{R}$ to $\curly{1,2,3,\ldots,k}$ such that

@@

The discretization edges $e_{1:k-1}$ mark the boundaries between discretization intervals.
In this work they are restricted to the midpoints between unique ascending instances of $\cX$.
Thus, each edge $e_i = \paren{\cu_{j} + \cu_{{j+1}}}/2$ for some $j$.
Two useful integer representations of $\Lambda_{\cX}$ can be written

\begin{equation}
\label{eq:disc_def}
  \Lambda_{\cX} = \anglebrackets{\lambda_1 < \lambda_2  < \ldots < \lambda_k} \equiv \anglebrackets{\gamma_1 , \gamma_2  , \ldots , \gamma_k}\text{,}
\end{equation}

@@

This section describes experiments conducted to evaluate the Bayesian discretization method.
All experiments were run on datasets from the publically available University of California, Irvine machine learning repository \citep{Lichman_2013}.
Variables are labeled alphabetically in the order given on the dataset information webpage.
In the figures that follow, shaded nodes correspond to initially discrete variables and the subscripts indicate the number of discrete instantiations.

Two experiments were conducted on each dataset.
The first experiment compares the performance of the Bayesian and MDL discretization methods on a known Bayesian network structure.
The structure was obtained by prediscretizing each continuous variable into $k$ uniform-width intervals, where $k$ is the median number of instantiations of the discrete variables, and using the structure with the highest Bayesian score from \num{1000} runs of the K2 algorithm with random variable orderings.
The second experiment compares the same methods applied when simultaneously discretizing and learning network structure.

The discretizations are compared using the mean cross validated log-likelihood of the data $D$ given the graph structure $G$ and discretization policies $\discset$.
Note that all log-likelihoods shown have been normalized by the number of data samples.
The log-likelihood has two components,

@@

The method for computing the MDL discretization policy is similar to the method for the Bayesian method.
For a Bayesian network with a single continuous variable $\cX$, the MDL objective function is

@@

The Bayesian and MDL discretization methods were tested on the Auto MPG data using the network shown in Figure~\ref{fig:auto_graph_1}.
This structure was obtained by prediscretizing each continuous variable into five uniform-width intervals, where five is the median cardinality of the discrete variables, and then taking the structure with highest likelihood from \num{1000} runs of K2.

@@

Table~\ref{table:auto_disc_table_1} lists the discretization edges and mean log-likelihoods under \num{10}-fold cross validation of the discrete Bayesian network resulting from each discretization method.
The MDL method does not produce any discretization edges, assigning one continuous interval to each continuous variable, and produces the structure with the lowest likelihood.
The cause behind MDL's discretization policy is discussed in Section~\ref{subsec:discuss_exp}.

\begin{table}[h]
  \centering
  \caption{
    Results from discretization of the Auto MPG dataset with fixed structure from Figure~\ref{fig:auto_graph_1}.
    The first five rows list the discretization edges for each continuous variable.
    The last row lists the mean cross-validated log-likelihood; positive values are better.
    The MDL method does not produce any discretization edges.
  }
  \input{table_autoMPG_1.tex}
  \label{table:auto_disc_table_1}
\end{table}

Figure~\ref{fig:auto_exp1_distr_1_3} shows the marginal probability density for variables $A$ and $C$ under the Bayesian discretization policy overlaid with the original Auto MPG data.
The resulting probability density is a good match to the original data.

\begin{figure}[ht]
  \centering
  \input{graph_auto_exp1_distr_1_3}
  \caption{
    Comparison of the Bayesian discretization policy for variables $A$ and $C$ to the original Auto MPG data learned with a fixed network.
    The marginal probability density closely matches the data.
  }
  \label{fig:auto_exp1_distr_1_3}
\end{figure}

@@

In this experiment the network structure was not fixed in advance and was learned simultaneously with the discretization policies.
Figure~\ref{fig:auto_graph_2} shows a learned Bayesian network structure and the corresponding numbers of intervals after discretization for each continuous variable.
This result was obtained by running Algorithm~\ref{alg:structure_learn} fifty times using the Bayesian method and choosing the structure with the highest likelihood.

\begin{figure}[ht]
  \centering
  \input{graph_autocar_2}
  \caption{The discrete-valued Bayesian network learned from the Auto MPG dataset using the Bayesian method.}
  \label{fig:auto_graph_2}
\end{figure}

Figure~\ref{fig:auto_exp2_distr_1_3} compares the Bayesian discretizaton policy for variables $A$ and $C$ in the learned network with the original Auto MPG data.
The color of a discretized region indicates the marginal probability density of a sample from $P(A,C)$ being drawn from that region.
Although there are fewer discretization edges for $A$ and $C$ in the learned network, the marginal distribution is still captured.
The discretization policy will vary according as the network structure changes, and it still produces high-quality discretizations.

\begin{figure}[ht]
  \centering
  \input{graph_auto_exp2_distr_1_3}
  \caption{
    Comparison of the Bayesian discretization policy for variables $A$ and $C$ to the original Auto MPG data learned simultaneously with the network structure.
    Although the number of discretization edges is less than those in Figure~\ref{fig:auto_exp1_distr_1_3}, the probability distribution still closely matches the original data.
  }
  \label{fig:auto_exp2_distr_1_3}
\end{figure}

@@

The Bayesian and MDL discretization methods were tested on the Wine data using the network shown in Figure~\ref{fig:wine_graph_1}.
This structure was obtained by prediscretizing each continuous variable into three uniform-width intervals, where three is the median cardinality of the discrete variables, and then taking the structure with highest likelihood from \num{1000} runs of K2.

Table~\ref{table:wine_disc_table_1} lists the discretization edges and mean log-likelihoods under \num{10}-fold cross validation of the Bayesian network resulting from each discretization method.
The Bayesian method outperforms the MDL method in likelihood by a significant margin.
The MDL method creates significantly fewer discretization edges.
% The heuristic method discretizes the variables $G$, $H$, and $J$ into one fewer interval and the variable $M$ into one more interval than the optimal Bayesian approach, resulting in a slightly lower mean cross validated log likelihood.
% The overall performance of the heuristic method is still very high, and if runtime is the primary concern, the heuristic method can be considered the better approach.
% \tim{Do we need to list runtime too?}

Some discretization edges appear in all three discretization methods, such as \num{1.42} and \num{2.35} for variable $C$ and \num{0.785} for variable $L$.
This indicates that MDL indeed can find some important discretization edges, but it is not sensitive enough to find more edges.

\begin{figure}[ht]
  \centering
  \scalebox{0.8}{\input{graph_wine_1}}
  \caption{Bayesian network structure obtained from running K2 on the prediscretized Wine dataset}
  \label{fig:wine_graph_1}
\end{figure}

\begin{table}
  \centering
  \caption{
    Results from discretization of the Wine MPG dataset with fixed structure.
    % The first thirteen rows list the discretization edges for each continuous variable.
    % The last row lists the mean cross-validated log-likelihood; positive values are better.
    % The optimal Bayesian approach has the best discretization result in terms of likelihood, but the heuristic method obtains a similar result with nearly the same likelihood.
    % The MDL method produces discretization edges for more than half of the continuous variables but does not produce enough discretization intervals.
    % Some discretization edges appear in the results of all three methods.
    Bold discretization edges were identified by both methods.
  }
  \scalebox{1.0}{
  \input{table_wine_1.tex}
  }
  \label{table:wine_disc_table_1}
\end{table}

Figure~\ref{fig:wine_exp1_distr} compares the Bayesian and MDL discretizaton policies for variables $E$ and $K$ with the original Wine data.
The discretization edges \num{17.9} for $E$ and \num{34.6} for $K$ appear in both plots.
The MDL method does not use enough intervals for discretization.
Relative sensitivities of each method to the input data is discussed in Section~\ref{subsec:discuss_exp}.

\begin{figure}[ht]
  \centering
  \input{graph_wine_exp1_distr}
  \caption{Comparison of the discretization policies for variables $A$ and $C$ obtained using the Bayesian and MDL methods}
  \label{fig:wine_exp1_distr}
\end{figure}

@@

Figure~\ref{fig:wine_exp2_distr} shows the discretization policy for variables $E$ and $K$ obtained with the Bayesian method.
The discretization edge also $E = 17.9$ appears in both discretization policies from the fixed network (Figure~\ref{fig:wine_exp1_distr}).
This suggests that discretization edges can be robust against network structure.
Furthermore, the discretization edge at $E = 23.5$ present in the fixed-structure case is missing in the learned structure.
This is caused by $E$ having twice as many parents in the learned network structure.
The more parents a variable has, the less discretization intervals it can to support, as the set of statistics required to define the resulting distribution increases exponentially with the number of parents.

\begin{figure}[ht]
  \centering
  \scalebox{0.7}{\input{graph_wine_2}}
  \caption{The discrete-valued Bayesian network learned from the Wine dataset using the Bayesian method.}
  \label{fig:wine_graph_2}
\end{figure}

\begin{figure}[ht]
  \centering
  \input{graph_wine_exp2_distr_5_11}
  \caption{
    The discretization policy for variables $E$ and $K$ on the learned network.
    The discretization edge $E = 17.9$ also appears in the policies for the fixed network.
  }
  \label{fig:wine_exp2_distr}
\end{figure}

@@

The Housing dataset contains variables related to the values of houses in Boston suburbs.
The dataset has \num{506} samples over fourteen variables.
Only variables $D$ and $I$ are discrete-valued, with \num{2} and \num{9} instantiations, respectively.
Despite their being continuous, several variables in the Housing dataset possess many repeated values.
The following experiments were conducted with a maximum of three parents per variable to prevent running out of memory when running MDL.

@@

The Bayesian network structure in Figure~\ref{fig:housing_graph_1} was obtained by prediscretizing each continuous variable into five uniform-width intervals and then running the K2 algorithm \num{1000} times and choosing the network with the highest likelihood.
Table~\ref{table:housing_disc_table_1} shows the numbers of intervals after discretization for each continuous variables and the log-likelihood of the dataset based on each discretization method.
MDL method does not produce any discretization edges for most variables.
% The heuristic Bayesian method produces far more discretization edges than the Bayesian method due to repeated instances in the dataset.
% The numbers of intervals returned by the each method illustrate the over-sensitive nature of the heuristic method and the under-sensitive nature of the MDL method.
% \tim{I am not sure we are using `sensitive' correctly. Sensitive to what? Perhaps `biased towards higher edge counts'?}
The relative weighting of each method's objective function will be discussed in Section~\ref{subsec:discuss_exp}.

% The heurstic method produces the discretization with the highest cross-validated log-likelihood; and although the heuristic method leads to a higher number of intervals, its likelihood is not significantly better than the Bayesian approach.
% This demonstrates that the Bayesian approach captures the most important discretization edges.

\begin{figure}[ht]
  \centering
  \scalebox{0.8}{\input{graph_housing_1}}
  \caption{Bayesian network structure obtained from running K2 on the prediscretized Housing dataset}
  \label{fig:housing_graph_1}
\end{figure}

\begin{table}
  \centering
  \caption{
    Discretization policy summary of the Housing dataset based on the fixed network structure shown.
    The first twelve rows show the numbers of intervals after discretization for each method.
    The last row is the mean cross-validated log-likelihood based on each discretization result.
    % The heuristic method is over-sensitive \todo{to what?} and discretizes some variables into too many intervals.
    % The MDL method is less sensitive and on many cases does not produce any discretization intervals.
    % The optimal Bayesian approach produces a reasonable number of discretization edges.
  }
  \input{table_housing_1.tex}
  \label{table:housing_disc_table_1}
\end{table}

Figures~\ref{fig:housing_exp1_distr_3_5} and \ref{fig:housing_exp1_distr_8_5} show the discretization policy learned using the Bayesian approach on the fixed network shown in Figure~\ref{fig:housing_graph_1}.
The scatter points in Figure~\ref{fig:housing_exp1_distr_3_5} were jittered to show the quantity of repeated values for variables $C$ and $E$.
Each repeated point forms a single discrete region, thereby encouraging discretization.
In contrast, the samples for $H$ are well spread out, resulting in fewer discretization regions and larger discretization intervals.

\begin{figure}[ht]
  \centering
  \input{graph_housing_exp1_distr_3_5}
  \caption{
    The discretization policies for variables $C$ and $E$ learned using the Bayesian approach on the fixed network.
    The scatter points were jittered to reveal the repeated values in the dataset.
    The repeated values lead to a high number of discretization edges, because repeated values can easily form discretized regions by their own.
  }
  \label{fig:housing_exp1_distr_3_5}
\end{figure}

@@

Figure~\ref{fig:housing_graph_2} shows a learned Bayesian network structure and the corresponding numbers of intervals after discretization for each continuous variable.
Variable $D$ is neighborless in both the learned and fixed networks.
The continuous variables $C$, $E$, $J$ and $K$, which are all connected through $C$, have many discretization intervals.
Typically, when discretizing a variable, the expected number of discretization intervals is close to the highest cardinality among variables in its Markov blanket.
This naturally leads to clusters of variables with many discretization intervals.

@@

\begin{figure}[ht]
  \centering
  \input{graph_housing_exp2_distr_3_5}
  \caption{
    The discretization policy for variables $C$ and $E$ on the learned network in Figure~\ref{fig:housing_graph_2}.
    The high number of discretization edges are caused by repeated values in $C$ and $E$.
  }
  \label{fig:housing_exp2_distr_3_5}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \input{graph_exp_discuss}
  \caption{A simple Bayesian network used to demonstrate the sensitivity of each method}
  \label{fig:exp_discuss}
\end{figure}

To quantatively discuss the sensitivity of each methods to the number of discretization intervals, consider the Bayesian network in Figure~\ref{fig:exp_discuss}, where $\cX$ is continuous and $P_1$ and $P_2$ are discrete.
If $\Pa_{\cX} = \curly{P_1, P_2}$ and $D_{\cX} = \curly{1.0,2.0,3.0, \ldots n}$, then the corresponding objective functions for the discretization methods are:

\begin{small}
  \begin{equation}
  \begin{aligned}
  f_{\text{MDL}} & = \overbrace{Z_1 \cdot k + \ln(k) + \ln {{n +k - 1}\choose{k -1}} }^{\text{Penalty Term}}+   \overbrace{\rule{0pt}{1.35em} \ln(N) \cdot I(X^*, \Pa_X)}^{\text{Edge Position Term}} \\
  f_{\text{Bayesian}} &= \underbrace{{\vrule width0pt height0pt depth1.48em\relax}  Z_2 \cdot k + \sum_{i=1}^k \ln {{\gamma_i + J_P - 1}\choose{J_P -1}}}_{\text{Penalty Term}} + \underbrace{\sum^k_{i=1} \ln \left(   {{\gamma_i !}\over{n^{(P)}_{i,1}! n^{(P)}_{i,2}! \cdots n^{(P)}_{i,J_P}!    }} \right) }_{\text{Edge Position Term}}
  % f_{\text{Heuristic}} &= \overbrace{Z_2 \cdot k + \sum_{i=1}^k \ln {{\gamma_i + J_{P_1} - 1}\choose{J_{P_1} -1}} + \sum_{i=1}^k \ln {{\gamma_i + J_{P_2} - 1}\choose{J_{P_2} -1}} }^{\text{Penalty Term}}+\\
  % & \overbrace{\sum^k_{i=1}  \left[ \ln \left(   {{\gamma_i !}\over{n^{(P_1)}_{i,1}! n^{(P_1)}_{i,2}! \cdots n^{(P_1)}_{i,J_{P_1}}    }} \right)  + \ln \left(   {{\gamma_i !}\over{n^{(P_2)}_{i,1}! n^{(P_2)}_{i,2} \cdots n^{(P_2)}_{i,J_{P_2}} }} \right) \right]}^{\text{Edge Position Term}}
  \end{aligned}\text{,}
  \end{equation}
\end{small}

\noindent
where $Z_1$ and $Z_2$ are constant over discretizations, $k$ is the number of discretization intervals, and $I\paren{A, B} = \sum_{a,b} \hat{P}(a,b) \ln\frac{\hat{P}(a,b)}{\hat{P}(a)\hat{P}(b)}$ is the mutual information based on estimated probabilities.
Note that the third term of $f_\text{MDL}$ had been approximated using $H(p)$ (see Equation~\ref{eqn:MDL}) in the original work by \cite{Friedman_1996}, but here is written without that approximation.

@@

% \section*{Appendix}
